---
title: '[Re] Local alignment statistics - simulating polypeptides'
author: "Nathan Brouwer"
date: "11/18/2019"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Polypeptide simulation

Simulation of polypeptides is done using R's sample() function.  This requires a vector of possible amino acids letters and their probabilities of occurring.  The following code expalins the general principles.

### Possible amino acids

Make a vector of all letters that represent amino acid
```{r}
# 1 letter codes 
aa1 <- c("A", "C", "D", "E", "F", "G", "H",
         "I", "K", "L", "M", "N", "P", "Q",
         "R", "S", "T", "V", "W", "Y")
```

### AA Frequencies

The frequencies below were derived from Robinson and Robinson.  For further details see the associated script.

```{r}
robinson.aafreq <- c(0.07804747,0.01924601,0.05363974,0.06294860,0.03855640, 
                     0.07377157,0.02199227,0.05141964,0.05743832,0.09019139,
                     0.02242519,0.04487258,0.05202795,0.04264360,0.05129531,
                     0.07119847,0.05841294,0.06440942,0.01329837,0.03216475)
```


### Select a single random amino acid

Example: Randomly select a single amino acid with equal frequency

```{r}

sample(x = aa1, size = 1, replace = T)

```


### Create a random polypeptide

 Make a vector of 20  amino acids (eg, a simulated polypeptide), asumming equal frequencies

```{r, echo = T}

pp.length <- 20
sample(x = aa1, size = pp.length, replace = T)

```

Make a vector of 20  amino acids (eg, a simulated polypeptide), asumming frequencies from Robinson and Robinson
```{r, echo = T}

pp.length <- 20
sample(x = aa1, size = pp.length, prob = robinson.aafreq, replace = T)

```


```{r}

```

